---
title: "Homework Two"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=F, message=F, echo = TRUE)
```
## Question one
We import the data and assign it to data
```{r}
library(tidyverse)
library(faraway)
data<-cheddar
```

We then make a regression model as such

```{r}
fit<-lm(taste~Acetic+H2S+Lactic, data= cheddar)
summary(fit)
```
as we can see, H2S and Lactic are the two variables that are significant at the 5% significance level

We then adjust Acetic and H2S to their original scales. 
```{r}
fitexp<- lm(taste ~ I(exp(1)^Acetic) + I(exp(1)^H2S)+Lactic, data=cheddar)

summary(fitexp)
```
As seen in the summary, only the variable lactic is significant at the 5% significance level. 

In ANOVA testing, the F test is used to compare nested models. Here, the models are not nested and thus we cannot use the F test. 
The models both give us $R^2$ values of 0.6518 and .0574 respectively. It is clear that the first model gives a better fit. 
To find the change in response, we must multiply the coefficient by the desired increase. In our case:

$$
0.01*3.9118=0.039188
$$

To find the change on H2S we use the following 

$$
e^{0.01}= 1.01
$$
meaning that there is a 101% change in H2S if there is an additive increase of .01 on the logarithmic scale. 

## Question Two

Importing the data

```{r}
data1<- teengamb
```

Creating the model

```{r}
gambmod <- lm(gamble ~ sex+status+income+verbal, data=data1)
summary(gambmod)
```

We are thus able to see that Sex and Income are significant variables at the 5% significance level. 

If we are to interpret sex, we know that sex is a binary variable. Thus, if beta = 1, then we know that the point is female, and that females have a negative relationship with gambeling in comparison to males. 

Creating a new model with just income

```{r}
gambmod1<- lm(gamble ~ income, data=data1)
summary(gambmod1)
```

We then use anova analysis to confuct the f test

```{r}
anova(gambmod, gambmod1)
```
We thus see with the pvalue of the f test that is it lower than our significance level. Thus we are able to conclude that the full model is better than the reduced model. 
