---
title: "Homework Two"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=F, message=F, echo = TRUE)
```

## Question One
We import our data and take a quick look to make sure it looks correct. 
```{r}
library(faraway)
library(car)
head(sat)
```

Now, we create a model with SAT as the response and expend ratio salary and takers as the predictor variables. 

```{r}
model1<-lm(total~expend+salary+ratio+takers, data = sat)

summary(model1)
```

### a

We plot the residuals to check for the constant variance assumption. 

```{r}
plot(fitted(model1), residuals(model1), xlab="Fitted", ylab="Residuals", abline(h=0))
```

Visually, we see that there is not any unusual patterns that indicate a non-constant variance. Regardless we can run a non-constant variance score test

```{r}
car::ncvTest(model1)
```

We see that the P value is greater that .05 and thus we can say that the variance of the residuals is constant. 

### b 

We can check the normality assumption by making a qq plot of the residuals

```{r}
qqnorm(residuals(model1))
qqline(residuals(model1))
```

Visually, we see that there is no cause to doubt that the normality assumption holds. Additionally, we can run a Shapiro test to see if the normality assumption holds. 

```{r}
shapiro.test(residuals(model1))
```

because of the sufficiently large P value, we can say that the normality assumption holds. 

### c 

we check the values of the diagonal hat matrix to see the leverage

```{r}
hat<- hatvalues(model1)
p=length(coefficients(model1))
n=length(fitted(model1))
ratio=p/n
plot(hat, main = 'Levarage plot', ylab = 'levarage', xlab = 'index')
abline(h=c(2,3)*ratio, col="blue", lty=2)
which.max(hat)
which.max(hat[hat!=max(hat)])
text(44, hat["Utah"], labels="UT", cex=.7, font=2)
text(5,hat["California"], labels="CA", cex=0.7, font=2)
```
we see that california and Utah are levarage points

### d 

We now check for out-liers and see that there are none to be found. 

```{r}
car::outlierTest(model1)
```

### e 

we use R plotting tools to check the influence points. We see that they are Utah and California. 

```{r}
car::influencePlot(model1, id.method="identify", main="influence plot", sub="Circle Size is poportional to cook's distance")
```

### f 
We now use R to plot the relationships with the other other predictors and response. 

```{r}
car::avPlots(model1)
```

We see that there are weak positive linear relationships with expend and salary. There are negative relationships with ratio and takers. Ratio however has a weak negative relationship while takers is a strong negative relationship. 
# Question two

```{r}
head(prostate)

model2<- lm(lpsa~., data=prostate)

summary(model2)
```

Checking Constance variance assumption

```{r}
plot(fitted(model2), residuals(model2), xlab="fitter", ylab = "residuals", abline(h=0))
ncvTest(model2)
```

we see that our P value is large and visually, there is nothing strange. Thus our assumption holds. 

#### b 

checking the normality assumption. 

```{r}
shapiro.test(residuals(model2))
qqnorm(residuals(model2))
qqline(residuals(model2))

```
P value is large, and there is nothing wrong visually, so the normality assumtion holds. 

```{r}
hat <- hatvalues(model2)
p <- length(coefficients(model2))
m<-length(fitted(model2))
ratio<-p/m

plot(hat,main = 'Hat Values')
abline(h=c(2,3)*ratio,col='blue', lty=2)
which.max(hat)
text(32,hat[32], labels=round(hat[32], 4), cex=0.7, font=2)
sum(hatvalues(model2))
```
and so, visually we see that there is one high leverage value that is affecting the model. which has a value of .3305 and is the 32th observation. the sum of hat values is nine so it checks out. 

#### d 
checking for outliers

```{r}
outlierTest(model2)
```
we see that there are no outliers. 

#### e 
checking for influence points 

```{r}
influencePlot(model2, id.method="identify", main="influence Plot", sub="Size is proportional to cook's distance")
```
we see that 32 has the influence observation. 

#### f

```{r}
avPlots(model2)
```
We see that age and lcp have negative linear relations, age being a strong relationship. all the others have positive relationships, except that lweight and lcavol have strong positive relationships. 
